---
title: "ETAN Database Exploration Tool"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = FALSE, cache.lazy = FALSE, autodep = TRUE, warning = FALSE, 
  message = FALSE, echo = TRUE, dpi = 180,
  fig.width = 8, fig.height = 5, echo = FALSE
  )
```

```{r include=FALSE}
#library set-up=================================================================
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
library(tidyverse)
library(data.table)
library(DT)
library(networkD3)
library(tidygraph)
library(ggraph)
library(readxl)
library(here)
library(scales)
library(gridExtra)
library(shiny)
```

```{r include=FALSE}
#data set-up====================================================================
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
data_file = "ssms_queries_20210610.xlsx"

data_file_date = data_file %>%  
  str_remove_all("[:punct:]") %>% 
  str_remove_all("[:alpha:]") %>%  
  lubridate::ymd()

data_file_path = data_file %>%
  paste0("analysis/", .) %>%
  here()

data_file_sheets = data_file_path %>% 
  excel_sheets()

query_list = data_file_sheets %>%  
  map(~ read_xlsx(data_file_path,
                  sheet = .x,
                  skip = 3) %>%  
        as.data.table) %>%  
  set_names(data_file_sheets)

count_list = query_list$table_record_count %>%  unique()
constraint_list = query_list$table_constraints
view_construction = query_list$view_construction
view_record_count = query_list$view_record_count
```

```{r}
cleaned_connections = 
  constraint_list %>%
  filter(str_detect(CONSTRAINT_NAME, "2")) %>%  #only keeps records where CONSTRAINT_NAME has a "2" in it
  separate(col = CONSTRAINT_NAME, sep ="_2_", into = c("from_complete", "to_complete")) %>%  
  na_if("NA") %>%  #makes sure string NAs are registered 
  filter(complete.cases(.)) %>%  #removes incomplete records
  filter(!str_detect(from_complete, "description_")) %>%  #removes records where from has "description_" - should be fixed later but don't know how to now
  filter(str_detect(to_complete, "_")) %>% #removes records that do not have schema_table
  separate(col = from_complete, sep ="_", into = c("type", "from_table", "from_column"), remove = F) %>% 
  separate(col = to_complete, sep ="_", into = c("to_schema", "to_table"), remove = F) %>%  
  mutate(from_complete = str_sub(from_complete, 4,-1), 
         from_schema = CONSTRAINT_SCHEMA, 
         from_complete = paste0(from_schema, "_", from_table)) %>%  
  merge(count_list[, c(1,3)], by.x = "from_table", by.y = "TableName") %>%  
  merge(count_list[, c(1,3)], by.x = "to_table", by.y = "TableName", suffixes = c("_from", "_to"), all.x = T) %>%
  select("CONSTRAINT_CATALOG", "type", contains("from"), contains("to")) %>%  
   rename(c('to_rows' = 'rows_to', 'from_rows' = 'rows_from')) %>% 
  arrange(desc(from_rows)) %>%  
  mutate(as_of = Sys.Date())


cleaned_connections_0filtered = cleaned_connections %>%  
  filter(to_rows > 0, 
         from_rows > 0)

cleaned_view_construct =  view_construction %>%  
  mutate(to_complete = str_glue("{schema_name}_{view_name}"), 
         from_complete = str_glue("{referenced_schema_name}_{referenced_entity_name}")) 
```

```{r}
#var defintion==================================================================
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
```

```{r}
#custom function defintion======================================================
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
source(here("code/helpers_networkPlots.R"))
source(here("code/utils_helpers_general.R"))
source(here("code/utils_helpers_DT.R"))
```


# Introduction/Summary

This tool was made to help users explor the ETAN database. Plots and tables have been provided to report descriptive statistics about tables and views in the database. The tables included in this tool can be copied and saved locally. 

The data reported used for this tool was queried on `r data_file_date`.

# Database Schemas

The below, left plot details the number of tables per each schema. The below, right plot details the total number of records in each schema - the coloring depicts the number of records for unique tables given a schema. Most records are in a handful of schemas and only from a few tables therein.

```{r}
plot_schema_tables = 
  count_list %>%  
  group_by(SchemaName) %>%  
  count(sort = T, name = "Tables") %>% 
  mutate(SchemaName = fct_reorder(SchemaName, Tables, sum)) %>%
  ggplot() + 
  geom_col(aes(SchemaName, Tables)) + 
  theme(legend.position = "none") + 
  coord_flip() + 
  labs(title = "Tables per Schema")

plot_schema_records = count_list %>%
  mutate(SchemaName = fct_reorder(SchemaName, rows, sum)) %>% 
  ggplot() + 
  geom_col(aes(SchemaName, log10(rows), fill = TableName)) + 
  theme(legend.position = "none") + 
  coord_flip() + 
  scale_fill_viridis_d() +
  labs(title = "Records per Schema")

grid.arrange(plot_schema_tables, plot_schema_records, ncol = 2)
```

 
# Database Tables
```{r}
count_list %>%  
  ggplot() + 
  geom_histogram(aes(rows), color = "black") +
  scale_x_continuous(trans='log10', 
                     breaks = trans_breaks("log10", function(x) 10^x),
                     labels = trans_format("log10", math_format(10^.x))) + 
  labs(title = "Histogram of Records per Table")
```

## ETAN Database Table Constraints

---

### Constraint Table

This table details all ETAN table constraintes - queried from ETAN database on `r Sys.Date()`.
```{r}
# checkboxGroupInput("column_limit", "Select columns to hide:", colnames(pretty_col_names(cleaned_connections)), inline = T)
# columns2hide = reactive({
#   match(input$column_limit, colnames(pretty_col_names(cleaned_connections)))
# })

cleaned_connections %>%
  select(-c('CONSTRAINT_CATALOG', "type", "as_of"), -contains("table")) %>% 
  dt_common(y = 500,
            pl = 10000,
            filter = "top",
            dom = "Bftipr") %>%
  DT::formatSignif('From Rows', 3) %>%
  DT::formatStyle(
    'From Rows',
    background = DT::styleColorBar(cleaned_connections$from_rows, 'lightblue'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  ) %>%
  DT::formatSignif('To Rows', 3) %>%
  DT::formatStyle(
    'To Rows',
    background = DT::styleColorBar(cleaned_connections$to_rows, 'lightblue'),
    backgroundSize = '100% 90%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'center'
  )

```

### Constraints Diagram

This diagram graphically depicts ETAN table contraints - it can be though of as an ERD (entity relationship diagram).
```{r}
tmp = cleaned_connections_0filtered

remove_list = c("info_InputSource", "info_InputUser", "Info_EnuInputLocation")

schema_remove = function(data, schema){
  data %>% 
    filter(!str_detect(to_complete, schema),
         !str_detect(from_complete, schema))
}

schema_keep_only = function(data, schema){
  data %>% 
    filter(str_detect(to_complete, schema),
           str_detect(from_complete, schema))
}
```

```{r}
checkboxGroupInput("from_schema", "Keep Only", inline = T,
                   choices = sort(unique(c(tmp$from_schema, tmp$to_schema))),
                   selected = sort(unique(c(tmp$from_schema, tmp$to_schema))))
checkboxGroupInput("to_schema", "Keep Only", inline = T,
                   choices = sort(unique(c(tmp$from_schema, tmp$to_schema))),
                   selected = sort(unique(c(tmp$from_schema, tmp$to_schema))))
# textInput("filter_1", "Keep Only")
yolo = reactive({input$filter_1})

observe({
  print(input$filter_1)
})

renderForceNetwork({
tmp %>%
  filter(str_detect(from_schema, input$from_schema %>%
                      paste(collapse  = "|"))) %>%
  filter(str_detect(to_schema, input$to_schema %>%  
                      paste(collapse  = "|"))) %>%
  make_force_network()
})
```

```{r}
cleaned_view_construct %>%  
  schema_remove("info_InputSource") %>% 
  schema_remove("info_InputUser") %>% 
  schema_remove("info_EnuInputLocation") %>% 
  schema_remove("config_") %>% 
  schema_remove("comm_") %>% 
  filter(str_detect(to_complete, "vehicl")) %>% 
  make_force_sankey()
```

## ETAN Database Views

---

This table depicts the views contained in the ETAN database. The _To Complete_ column contains full view names (schema+view) and _From Complete_ column contains full table names for the table which make said view. The table is orgainzed in a long/machine readable format.
```{r}
cleaned_view_construct %>%  
  select(entity_type, from_complete, to_complete) %>% 
  dt_common(y = 500,
            pl = 10000,
            filter = "top",
            dom = "Bftipr")
```








